Open Source, Collaborative Curriculum

### 2017-04-27 siip.tex
subsections rather than paragraphs

          \newcommand{\authorname}{Kathryn~D.~Huff }
          \newcommand{\authoremail}{katyhuff@illinois.edu}
          \newcommand{\authorsite}{arfc.npre.illinois.edu}

@@

          \pagestyle{fancy}
          %\pagenumbering{gobble}
          %\fancyhead[location]{text}
          % Leave Left and Right Header empty.
          %\lhead{}
          %\rhead{}
          \lhead{\textcolor{gray}{PI: \authorname (NPRE)\\\authoremail}}
          \rhead{\textcolor{gray}{Collaborative Open Source Curriculum Development\\}}
          %\rhead{\textcolor{gray}{\thepage/\totalpages{}}}
          \renewcommand{\headrulewidth}{0pt}
          \renewcommand{\footrulewidth}{0pt}
          \fancyfoot[C]{\footnotesize \textcolor{gray}{\authorsite}}

          \section*{Problem}
          At this very moment, hundreds of professors in the United States are 
          simultaneously preparing lessons on transposing a matrix.
          They are doing so largely without receiving feedback from one another 
          or directly building on one another's experience 
          \cite{green_building_2014}. In this way, 
          professors spend an enormous amount of time duplicating curriculum 
          development efforts already tackled by colleagues. What's worse is 
          that these efforts are rarely, if ever, reviewed by, shared with, or 
          extended upon by peers.

@@

          \section*{Proposed Solution}
          We propose a small-scale proof-of-concept for collaborative, open 
          source, curriculum development to improve the transfer of lessons 
          learned between instructors of the same course (either at a single 
          university or among different campuses). This prototype collaboration 
          will provide a template which could be adopted for collaboration 
          among faculty teaching courses with an inherently larger scale (e.g. 
          CS101)

          Faculty in Nuclear Engineering from four institutions\footnote{Paul 
          Wilson at (UW Madison), Steven Skutnik (UT Knoxville), Jeremy Roberts 
          (Kansas State University), Anthony Scopatz (University of South 
          Carolina)} have agreed to be participants in 
          this prototyping effort. We all currently use 
          \href{https://github.com}{GitHub} to store, 
          revise, and collaborate on research, especially source code. 
          Additionally, this select group already started to host our course 
          curricula online as well, but these are typically single author 
          repositories (e.g. \cite{huff_npre412_2017}).

@@

          \paragraph{In a typical open source project,} a main copy of the 
          repository (or ``fork'') holds the 
          official copy of the software package. Individual developers each have their own 
          forks where they can work on features and 
          bug fixes. When the developer makes changes that are ready for prime 
          time, they make a ``pull request'' to the main fork, it is 
          reviewed by their collaborators, and it is eventually merged into the 
          main fork where it can be used by all. Figures 
          \ref{fig:sub1} and \ref{fig:sub2} show how the workflow used for many 
          open source projects can be leveraged toward course development.

          Development of this proof-of-concept curriculum will be undertaken 
          similarly to an open source software project.

          Some features of open source development that will be emulated 
          include those that cultivate collaboration. In open source, 
          for example, new code features are described in tickets and discussed 
          before implementation. Small, atomic changes are reviewed before 
          merging. These discussions and incremental reviews nurture 
          communities of practice.

@@

          \paragraph{A two-day kick-off workshop (\$15K)} at UIUC will allow the 
          participants to sketch the initial framework of 
          the collaboration and coordinate logistics. The workshop will 
          deliver:

@@

          Invited participants in the first workshop will include:
          \begin{itemize}
                  \item UIUC PI Huff and UIUC Co-PI Davis
                  \item external collaborating Professors Wilson, Roberts, 
                          Skutnik, and Scopatz
                  \item up to 5 UIUC professors who may be likely to extend 
                          this work in the future for their own courses.
          \end{itemize}

          \paragraph{Between Workshops} work will commence on developing 
          lessons. These lessons will be submitted by pull request to the main 
          repository and each lesson will include:
          \begin{itemize} 
                  \item associated learning objectives (identified previously)
                \item content (e.g. speaking notes, presentation material, 
                        derivations, worked examples, active learning  
                          exercises, external readings, videos, images)
                  \item learning assessments (e.g. project descriptions, exam questions) 
          \end{itemize}

@@

          \paragraph{A two-day retrospective workshop (\$15K)} at UIUC will wrap up the 
          project. Discussion concerning the process will allow reflection as 
          well as a collection of lessons learned.
          Experiences live testing the developed course material (for example 
          in NPRE412 at UIUC in Fall 2017) will be shared and suggested 
          improvements will be captured as feature requests on GitHub.  
          The lessons learned from the year-long process will be captured:

@@

          \section*{Potential Impact}
          This work will have immediate student and faculty outcomes at UIUC, 
          USC, KSU, UT Knoxville, and UW Madison. It will additionally have long term 
          student and faculty outcomes at UIUC. Beyond this, it may have 
          institutional outcomes at UIUC.

@@

          \section*{Timeline and Deliverables}

@@

          \section*{Budget Justification}
          The two workshops will 
          Explain travel for external collaborators, meals, meeting space, etc. 
          Host it in the Union?

          \section*{Departmental Support}
          The Department of Nuclear, Plasma, and Radiological Engineering will 
          support this work in kind with release time for Kathryn Huff through 
          her Start-Up funds.  Additionally, NPRE will support workshop activities 
          with administrative effort and by providing space. Finally, 
          collaborating external participants will contribute in kind with 
          summer time hours.

>>>>

          \newcommand{\authorname}{Kathryn~D.~Huff }
          \newcommand{\authoremail}{kdhuff@illinois.edu}
          \newcommand{\authorsite}{arfc.npre.illinois.edu}

@@

          \pagestyle{fancy}
          %\pagenumbering{gobble}
          %\fancyhead[location]{text}
          % Leave Left and Right Header empty.
          %\lhead{}
          %\rhead{}
          \lhead{\textcolor{gray}{PI: \authorname (NPRE)\\\authoremail}}
          \rhead{\textcolor{gray}{Collaborative Open Source Curriculum Development\\}}
          %\rhead{\textcolor{gray}{\thepage/\totalpages{}}}
          \renewcommand{\headrulewidth}{0pt}
          \renewcommand{\footrulewidth}{0pt}
          %\fancyfoot[C]{\footnotesize \textcolor{gray}{\authorsite}}

          \section{Problem}
          At this very moment, hundreds of professors in the United States are 
          simultaneously preparing lessons on transposing a matrix.
          They are doing so largely without receiving feedback from one another 
          or directly building on one another's experience 
          \cite{green_building_2014}. In this way, 
          professors spend an enormous amount of time duplicating curriculum 
          development efforts already tackled by colleagues. What's worse is 
          that these efforts are rarely, if ever, reviewed by, shared with, or 
          extended upon by peers.

@@

          \section{Proposed Solution}
          We propose a small-scale proof-of-concept for collaborative, open 
          source, curriculum development to improve the transfer of lessons 
          learned between instructors of the same course (either at a single 
          university or among different campuses). This prototype collaboration 
          will provide a template which could be adopted for collaboration 
          among faculty teaching courses with an inherently larger scale (e.g. 
          CS101)

          Faculty in Nuclear Engineering from four institutions\footnote{Paul 
          Wilson at (UW Madison), Steven Skutnik (UT Knoxville), Jeremy Roberts 
          (Kansas State University), Anthony Scopatz (University of South 
          Carolina), and Robert Borrelli (University of Idaho)} have agreed to be participants in 
          this prototyping effort. We all currently use 
          \href{https://github.com}{GitHub} to store, 
          revise, and collaborate on research, especially source code. 
          Additionally, this select group already started to host our course 
          curricula online as well, but these are typically single author 
          repositories (e.g. \cite{huff_npre412_2017}).

@@

          \subsection{Open Source Workflow}
          Development of this proof-of-concept curriculum will be undertaken 
          similarly to an open source software project. 
          In particular, this work will emulate features of open source software development 
          that cultivate distributed collaboration. In open source, for 
          example, new code features are described in \emph{issues} or 
          \emph{tickets} and discussed before implementation. Small, atomic 
          changes are reviewed before they are merged into the 
          \emph{repository}. These discussions and incremental reviews nurture 
          communities of practice.

          In a typical open source project, a main copy of the 
          repository (or main \emph{fork}) holds the 
          official copy of the software package. Individual developers each have their own 
          \emph{forks} where they can work on features and 
          bug fixes. When the developer makes changes that are ready for prime 
          time, they make a ``pull request'' to the main \emph{fork}. That 
          \emph{pull request} is 
          reviewed by their collaborators, and it is eventually merged into the 
          main \emph{fork} where it can be used by all. Figures 
          \ref{fig:sub1} and \ref{fig:sub2} show how this open source software 
          development workflow
          can be leveraged toward course development.

@@

          \subsection{Kick-Off Workshop}
          A two-day kick-off workshop (\$15K) at UIUC will allow the 
          participants to sketch the initial framework of 
          the collaboration and coordinate logistics. The workshop will 
          deliver:

@@

          Invited participants in the first workshop will include:
          \begin{itemize}
                  \item UIUC PI Huff and UIUC Co-PI Davis
                  \item external collaborating Professors Wilson, Roberts, 
                          Borrelli, Skutnik, and Scopatz
                  \item up to 5 UIUC professors who may be likely to extend 
                          this work in the future for their own courses.
          \end{itemize}

          \subsection{Interim Period}
          Between Workshops, work will commence on developing 
          lessons. These lessons will be submitted by pull request to the main 
          repository and each lesson will include:
          \begin{itemize} 
                  \item associated learning objectives (identified previously)
                \item content (e.g. speaking notes, presentation material, 
                        derivations, worked examples, active learning  
                          exercises, external readings, videos, images)
                  \item learning assessments (e.g. project descriptions, exam questions) 
          \end{itemize}

@@

          \subsection{Retrospective Workshop}
          A two-day retrospective workshop (\$15K) at UIUC will wrap up the 
          project. Discussion concerning the process will allow reflection as 
          well as a collection of lessons learned.
          Experiences live testing the developed course material (for example 
          in NPRE412 at UIUC in Fall 2017) will be shared and suggested 
          improvements will be captured as feature requests on GitHub.  
          The lessons learned from the year-long process will be captured:

@@

          \section{Potential Impact}
          This work will have immediate student and faculty outcomes at UIUC, 
          USC, KSU, UT Knoxville, and UW Madison. It will additionally have long term 
          student and faculty outcomes at UIUC. Beyond this, it may have 
          institutional outcomes at UIUC.

@@

          \section{Timeline and Deliverables}

@@

          \section{Budget Justification}
          The two workshops will 
          Explain travel for external collaborators, meals, meeting space, etc. 
          Host it in the Union?

          \section{Departmental Support}
          The Department of Nuclear, Plasma, and Radiological Engineering will 
          support this work in kind with release time for Kathryn Huff through 
          her Start-Up funds.  Additionally, NPRE will support workshop activities 
          with administrative effort and by providing space. Finally, 
          collaborating external participants will contribute in kind with 
          summer time hours.
